%*******************************************************
% Abstract
%*******************************************************
%\renewcommand{\abstractname}{Abstract}
\pdfbookmark[1]{Abstract}{Abstract}
\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Abstract}

Many computations running on high-performing systems does not make use
of the performance available. This problem is related to computations
failing to achieve strong scaling.

Copernicus is a system for execution of large-scale sampling tasks in
high-performance environments. It aims to achieve strong scaling,
regardless of underlying architecture. The system was originally
developed to run large scale bio-molecular simulations. However,
lacking a user-friendly way of describing computational projects, the
developers felt a need for an intuitive text-based input for
Copernicus.

<<<<<<< HEAD
This master's thesis describes a design and implementation of a
domain-specific language to meet the need of a suitable input
description for Copernicus. The language design is simple but still
manages to capture the abstract model of how a computational project
is executed. As there were no resembling solutions to the problem, the
language is inspired by elements from functional programming and
data-flow languages.

\highlight{SOME CONCLUSIONS}
=======
This master's thesis describes a design and implementation of a domain
specific language to meet the need of a better input for
Copernicus. The language design is simple but still can represent the
desired computational projects. It has both a flow-based and
functional aspect which makes it a powerful descriptive language.
>>>>>>> c3ff7e254160fea5f78b53377c3b0c2c63c93278

\endgroup			

\vfill
